---
title: "Start working with German Credit data"
format: html
editor: visual
---

To start working with the data, you might

-   download data on your computer,

-   change the path to data on your computer in the `load` function

-   render this file and look at the output in your default browser.

Continue working with the data by adding new code chunks, writing and executing the written code, rendering the file.

```{r, load libraries}
#| include: TRUE
#| echo: TRUE
# see more about chunk options, such as include and echo,  at https://quarto.org/docs/computations/execution-options.html
library(tidyverse)
```

To load the data in R environment, download the data from Canvas to your computer and use the command `read.csv` as below. Do not forget to substitute the path to the file to the path to the file on YOUR computer

```{r load data}
# Do not forget to substitute the path to the file to the path to the file on YOUR computer
# on Windows, the paths might look like  "C:\\Users\\YourUsername\\Documents\\data.csv"
# on MAC, the paths look like "/Users/YourUsername/Documents/data.csv"
data_credit <- read.csv("/Users/gote0002/Documents/GitHub/Business-Analytics/Data sets/German_credit/GermanCredit.csv") 
```

Take a look at the first couple of rows of the data:

```{r, first 6 rows}
head(data_credit)
```

Get a glimpse of the data:

```{r}
glimpse(data_credit)
```

Continue working with the data by adding new code chunks via menu Code/Insert chunk, writing the code, executing the written code or rendering the file.

```{r, working with data}
```
