---
title: "Start working with bike rental data"
format: html
editor: visual
---

To start working with the data, you might

-   download data file to your computer and to THE SAME FOLDER AS THIS QUARTO FILE (!!!!)

-   render this file and look at the output in your default browser.

Continue working with the data by adding new code chunks, writing and executing the written code, rendering the file.

```{r, load libraries}
#| include: TRUE
#| echo: TRUE
# see more about chunk options, such as include and echo,  at https://quarto.org/docs/computations/execution-options.html
library(tidyverse)
```

To load the data into R environment, download the data from Canvas to your computer\
and THE SAME FOLDER AS THIS QUARTO FILE and use the command `load` as below. Otherwise, you have to specify the full path to the data file in the `load` command.

```{r load bike data}
load("bike.RData") 
```

Take a look at the first couple of rows of the data:

```{r, first 6 rows}
head(bike)
```

Get a glimpse of the data:

```{r}
glimpse(bike)
```

Continue working with the data by adding new code chunks via menu Code/Insert chunk, writing the code, executing the written code or rendering the file.

```{r, working with data}
```
