---
title: "R functions used in week 1 and 2"
output: html_document
date: ""
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading R packages

```{r}
library("tidyverse")
library("openintro")
```

One can see the dimensions of data frame `arbuthnot` as well as the names of the variables and the first few observations by inserting the name of the dataset into the `glimpse()` function, as seen below:

```{r}
glimpse(arbuthnot)
```

We can access the data in a single column of a data frame by extracting the column with a \$.

```{r}
arbuthnot$boys
```

To learn what a function does and how to use it (e.g. the function’s arguments), just type in a question mark followed by the name of the function that you’re interested in into the console

```{r}
?ggplot
```

This opens a help file in R studio.

Adding a new variable to a data frame.

We are interested in using this new vector of the total number of baptisms to generate some plots, so we’ll want to save it as a permanent column in our data frame. We can do this using the following code:

```{r}
arbuthnot <- arbuthnot %>%
  mutate(total = boys + girls)
```

```{r}

```
