---
title: "Introduction to data: 1 - Language of data"
author: "Tetiana Gorbach"
date: ""
output: openintro::lab_report
---



# Welcome

Hello, and welcome to Introduction to data!

In this tutorial we will take you through concepts and R code that are essential for getting started with data analysis.


# Packages

Packages are the fundamental units of reproducible R code. They include reusable functions, the documentation that describes how to use them, and sample data. In this lesson we will make use of two packages:

 - tidyverse: Tidyverse is a collection of R packages for data science that adhere to a common philosophy of data and R programming syntax, and are designed to work together naturally. You can learn more about tidyverse here. But no need to go digging through the package documentation, we will walk you through what you need to know about these packages as they become relevant.
 - openintro: The openintro package contains datasets used in openintro resources. You can find out more about the package here.
Once we have installed the packages, we use the library() function to load packages into R.

Let’s load these two packages to be used in the remainder of this lesson.

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
```

# Data in R

In this lesson, we’ll begin by introducing the terminology of datasets and data frames in R.

One of the datasets that we will work with in this tutorial comes from the High School and Beyond Survey, which is a survey conducted on high school seniors by the National Center of Education Statistics.

The data are organized in what we call a data frame, where each row represents an observation or a case and each column represents a variable. If you ever use spreadsheets, such as a Google sheet or Excel, this representation should be familiar to you.

In this lesson we’ll work with the High School and Beyond dataset, stored in the openintro package. The data are stored in a data frame called hsb2. You can read more about this dataset here. Below is a preview of the dataset. You can use the arrow on the right hand side to scroll through the variables, and the numbers on the bottom to see different rows in the dataset

## Loading data into R
There are many ways of loading data into R depending on where your data are stored. In this lesson we’re using a dataset that is included in an R package so we can access this dataset by loading the package with the library() function. Other commonly used formats of data are plain text, comma separated values (CSV), Excel files (XLS or XLSX), or RData (the file format R uses to store data).

```{r, load data}
read_csv("/Users/gote0002/Documents/GitHub/Business Analytics/Data_sets/bike_sharing.csv")
```


In many of the tutorials in these series we will use data from existing R packages, so you won’t need to worry about the myriad ways of loading data into R. However, learning more about loading data into R is very important for when you’re working on an analysis with your own data. A resource we recommend for learning more about importing data into R is the Data Import chapter in R 4 Data Science by Grolemund and Wickham.

## Take a peek
When you want to work with data in R, a good first step is to take a peek at what the data look like. The glimpse() function is one good way of doing this. Click on the blue “Run Code” button to run the code below, and take a look at the output of the glimpse() function.

