---
title: "Week 1 Lesson 2"
format: html
editor: visual
---

# Exercise 1.

```{r, load libraries}
#| include: TRUE
#| echo: TRUE
# see more about chunk options, such as include and echo,  at https://quarto.org/docs/computations/execution-options.html
library(tidyverse)
```

To load the data in R environment, download the data from Canvas to your computer and use the command `load` as below. Do not forget to substitute the path to the file to the path to the file on YOUR computer

```{r load bike data}
# Do not forget to substitute the path to the file to the path to the file on YOUR computer
load("/Users/gote0002/Documents/GitHub/Business-Analytics/Data sets/Bike rental/bike.Rdata") 
```

Take a look at the first couple of rows of the data:

```{r, first 6 rows}
head(bike)
```

Get a glimpse of the data:

```{r}
glimpse(bike)
```

```{r, count types of weather}
bike%>%
  summarise(rented_member = sum(member))
sum(bike$member)
```

```{r}
bike%>%
  mutate(percentage_rented_member = member/total)
```

```{r, visualize number of days per type of weather}
bike%>%
  ggplot() + 
  geom_point(aes(x = instant, y = member, col = season)) + 
  theme_minimal()
bike%>%
  ggplot() + 
  geom_point(aes(x = instant, y = nonmember, col = season)) + 
  theme_minimal()

```

### Exercise 2.

First, remove all objects from the environment to start with the new exercise.

```{r}
rm(list = ls()) # remove all objects from the environment to start with the new exercise.
```

```{r}
h1 <- readRDS("/Users/gote0002/Documents/GitHub/Business-Analytics/Data sets/Hotel_demand/data_hotel/H1.rds") 
h2 <- readRDS("/Users/gote0002/Documents/GitHub/Business-Analytics/Data sets/Hotel_demand/data_hotel/H2.rds") 
```
