---
title: "Exercise to prepare for Lesson 2"
author: "Author Name"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
```

Comics data is not easily accessible with opentinro package!


Materials in this set of exercises is adjusted from https://openintrostat.github.io/ims-tutorials/02-explore/
# Exercise 1: Vizualizing categorical data
 - Graphical representation of two categorical variables
Side-by-side bar charts
Stacked bar charts
To normalize or not to normalize
 - Tabular representation of two categorical variables
Computation of margins
Counts vs proportions
Law of total probability
 - Graphical representation of one categorical variable
    --Marginal vs conditional
    --Bar chart
        --ordering
      ---data integrity check for levels
Pie chart

Exploring categorical data
In this tutorial we’ll learn about summarizing and visualizing data. Sometimes, we refer to the techniques covered in this tutorial as exploratory data analysis or EDA.

EDA is all about learning the structure of a dataset through a series of numerical and graphical techniques. When you do EDA, you’ll look for both general trends and interesting outliers in your data. You’ll also generate questions that will help inform subsequent analysis.

The emphasis of the first two lessons is on visualization: how to create and interpret graphical displays of your data. The third lesson focuses on numerical summaries of distributions and utilizes a simple, but powerful coding paradigm in R. We end with a case study that will allow you to synthesize these concepts and apply them to a more complex dataset that contains information on the characteristics of email that are associated with it being spam.

Before beginning the tutorial, we recommend you have some exposure to the R language and that you’ve taken the Introduction to data tutorial in this series.

Let’s get started.

Introduction
In this tutorial, you’ll be exploring data from a wide range of contexts. The first dataset comes from comic books. Two publishers, Marvel and DC, have created a host of superheroes that have made their way into popular culture. You’re probably familiar with Batman and Spiderman, but what about Mor the Mighty?

The comics dataset has information on all comic characters that have been introduced by DC and Marvel. Let’s take take a look at the dataset and explore it.
```{r load data}
glimpse(comics)
```

At the top the output tell us the dimensions of this dataset: over 15,000 cases and 11 variables. With one exception (appearances) we see that that all variables are characters, or of the chr data type. These are variables that we could potentially consider categorical, as they all have a finite number of levels.

The first case in the dataset is Peter Parker, alias: Spiderman, where his alias and character name are included in the first variable name. The second column, id, shows that Peter Parker’s personal identity is kept secret, and the third column tell us that his alignment is good; that he’s a superhero, not a super villain.

We can learn the different values of a particular character variable by using the distinct() function. It’s clear that the alignment variable can be “good” or “neutral”, but what other values are possible? Here, we pipe the comics data into the distinct() function, and then insert the align column into the distinct() function.
```{r}
comics%>% 
  distinct(align)

```

